### 2021-11-18 thesis/tex/oscillations.tex
update

What form does this new matrix $U$ take? By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three mixing angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure. We are now down to four degrees of freedom in the three dimensional case: three mixing angles
%TODO: Look CKM wikipedia, good explanation. Mention that it is not guaranteed that the mixing matrix that we only have one phase. We could Have three due to Dirac/majorana. Majorana phases are not relevant for oscillations because oscillations are not violating lepton number, but include them in PMNS.
of the form $\sin{(\theta_{ij})}/\cos{(\theta_{ij})} = s_{ij}/c_{ij}$, and one phase of the form $e^{i\delta}$. The mixing angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates that we allowed to mix. More on this in Section~\ref{sec:prop}.

We construct the PMNS matrix using a standard parametrization as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12} \nonumber \\
      & = 
    \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix} \nonumber \\
 %TODO: add dcp
 &=\begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13} & c_{12} c_{23}-s_{12} s_{23} s_{13} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}  & -c_{12} s_{23}-s_{12} c_{23} s_{13} & c_{23} c_{13}
    \end{pmatrix} \nonumber \\
&= \begin{pmatrix}
  U_{e 1} & U_{e2} & U_{e3} \\
   U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
   U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
\end{pmatrix}\,.
\end{align}

@@

Our treatment of the interaction Hamiltonian is similar except for the fact that its eigenstates lie in the flavor basis, conveniently allowing us
to letting it act directly on the flavor eigenstates:
\begin{align}
    \bra{\nu_\b}H_I &= V_\beta \bra{\nu_\b} \nonumber \\
                    &= \delta_{\b\eta} V_\beta \bra{\nu_\eta}\,.
\end{align}
Using Eq.~\ref{eq:V_matrix} in component form, we rewrite this as
\begin{align}\label{eq:V_rewrite}
    \delta_{\b\eta} V_\beta \bra{\nu_\eta} &= \delta_{\b\eta} (V_{CC}\delta_{\beta e} + V_{NC}) \bra{\nu_\eta} \nonumber \\
                                           &= V_{CC}\delta_{\b\eta}\delta_{\beta e} \bra{\nu_\eta} + V_{NC}\bra{\nu_\beta} \nonumber \\
    \implies \bra{\nu_\b}H_I\ket{\nu_\a}   &= V_{CC}\delta_{\b\eta}\delta_{\beta e} \bra{\nu_\eta}\ket{\nu_\a} + V_{NC}\bra{\nu_\beta}\ket{\nu_\a} \nonumber \\
                                           &= V_{CC}\delta_{\b\eta}\delta_{\beta e} \psi_{\a\eta} + V_{NC}\psi_{\a\b}
\end{align}
Now, combining Eq.~\ref{eq:t1} and Eq.~\ref{eq:V_rewrite}, we have for the full Hamiltonian
\begin{align}
    \bra{\nu_\b}H\ket{\nu_\a (x)} &= \left(p + \frac{m^2_1}{2E} + V_{NC}\right) \psi_{\a\b}(x) + \sum_\eta\sum_k \left(U_{\b k} \frac{\dm_{k1}}{2E} U^*_{\eta k} + V_{CC}\delta_{\b\eta}\delta_{\eta e} \right) \psi_{\a\eta}(x)
\end{align}
In this form, we see that the term $p + \frac{m^2_1}{2E} + V_{NC}$  is a common term to all flavor states, and does not affect the probability. It can be rotated away.
Thus
\begin{align}\label{eq:components}
    \bra{\nu_\b}H\ket{\nu_\a (x)} &= \sum_\eta\sum_k \left(U_{\b k} \frac{\dm_{k1}}{2E} U^*_{\eta k} + V_{CC}\delta_{\b\eta}\delta_{\eta e}\right) \psi_{\a\eta}(x) \nonumber \\
                                  &= i \dv{}{x}\psi_{\a\b}(x)\,.
\end{align}
If we form the vector 
\begin{align}
    \Psi_\a = \begin{pmatrix}
        \psi_{\a e} \\
        \psi_{\a \mu} \\
        \psi_{\a \tau}
    \end{pmatrix}\,,
\end{align}
we can write the Schrödinger equation on matrix form ($i \dv{}{x}\Psi_\a = H_F \Psi_\a$) and compare it with Eq.~\ref{eq:components} to see that the flavor Hamiltonian takes the form 
\begin{align}\label{eq:H_3gen}
    H_F &= \frac{1}{2E}(U M^2 U^\dagger + A) \nonumber \\
        &= \frac{1}{2E}\left[U \begin{pmatrix}
            0 & 0 & 0 \\
            0 & \dm_{21} & 0 \\
            0 & 0 & \dm_{31}
        \end{pmatrix} U^\dagger\right] + \sqrt{2}G_F N_e \begin{pmatrix}
            1 & 0 & 0 \\
            0 & 0 & 0 \\
            0 & 0 & 0
        \end{pmatrix}\,. 
\end{align}
This is the three-flavor neutrino oscillation Hamiltonian that we will solve numerically to obtain the evolution of $\Psi_\a$, whose squared components
are the probabilities
\begin{align}
    P_\alpha = \abs{\Psi_a}^2 &=\begin{pmatrix}
        \abs{\psi_{\a e}}^2 \\
        \abs{\psi_{\a \mu}}^{2} \\
        \abs{\psi_{\a \tau}}^{2}
    \end{pmatrix} \nonumber \\
    &=\begin{pmatrix}
        P_{\a e} \\
        P_{\a \mu} \\
        P_{\a \tau}
    \end{pmatrix} 
\end{align}

@@

\subsection{Two Neutrino Picture Solution}
By working in the limit where one of the neutrino masses can be neglected, 
we can reduce the complexity of the neutrino oscillations and obtain tractable analytical 
expressions. 
We now only consider the mixing between two neutrino flavors, say between $\ne, \nm$, and $\nu_1, \nu_2$. In other words, we let $\dm[31] = 0$.

>>>>

We construct the PMNS matrix by parametrizing it as as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12} \nonumber \\
     &= \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
         U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
         U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
      \end{pmatrix} \nonumber \\
      & = \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix} \nonumber \\
 &=\begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{12} c_{23}-s_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}}  & -c_{12} s_{23}-s_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{23} c_{13}
    \end{pmatrix}\,,
\end{align}
where $s_{ij} = \sin(\theta_{ij})$, $c_{ij} = \cos(\theta_{ij})$, $0 \leq \theta_{ij} \leq \pi/2$ and $ 0 \leq \delta_\mathrm{CP} \leq 2\pi$

By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure. We are now down to four degrees of freedom in the three dimensional case: three angles
%TODO: Look CKM wikipedia, good explanation. Mention that it is not guaranteed that the mixing matrix that we only have one phase. We could Have three due to Dirac/majorana. Majorana phases are not relevant for oscillations because oscillations are not violating lepton number, but include them in PMNS.
of the form $s_{ij}, c_{ij}$, and one phase of the form $e^{i\delta}$. The angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. Thus we refer to $\theta_{ij}$ as the mixing angles. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates $\nu_i$ that we allowed to mix. More on this in Section~\ref{sec:prop}.

@@

Our treatment of the interaction Hamiltonian is similar except for the fact that its eigenstates lie in the flavor basis, conveniently allowing us
to letting it act directly on the flavor eigenstates:
\begin{align}
    \bra{\nu_\b}H_I &= V_\beta \bra{\nu_\b} \nonumber \\
                    &= \delta_{\b\eta} V_\beta \bra{\nu_\eta}\,.
\end{align}
Using Eq.~\ref{eq:V_matrix} in component form, we rewrite this as
\begin{align}\label{eq:V_rewrite}
    \delta_{\b\eta} V_\beta \bra{\nu_\eta} &= \delta_{\b\eta} (V_{CC}\delta_{\beta e} + V_{NC}) \bra{\nu_\eta} \nonumber \\
                                           &= V_{CC}\delta_{\b\eta}\delta_{\beta e} \bra{\nu_\eta} + V_{NC}\bra{\nu_\beta} \nonumber \\
    \implies \bra{\nu_\b}H_I\ket{\nu_\a}   &= V_{CC}\delta_{\b\eta}\delta_{\beta e} \bra{\nu_\eta}\ket{\nu_\a} + V_{NC}\bra{\nu_\beta}\ket{\nu_\a} \nonumber \\
                                           &= V_{CC}\delta_{\b\eta}\delta_{\beta e} \psi_{\a\eta} + V_{NC}\psi_{\a\b}
\end{align}
Now, combining Eq.~\ref{eq:t1} and Eq.~\ref{eq:V_rewrite}, we have for the full Hamiltonian
\begin{align}
    \bra{\nu_\b}H\ket{\nu_\a (x)} &= \left(p + \frac{m^2_1}{2E} + V_{NC}\right) \psi_{\a\b}(x) \nonumber \\
                                  &+ \sum_\eta\sum_k \left(U_{\b k} \frac{\dm_{k1}}{2E} U^*_{\eta k} + V_{CC}\delta_{\b\eta}\delta_{\eta e} \right) \psi_{\a\eta}(x)
\end{align}
In this form, we see that the term $p + \frac{m^2_1}{2E} + V_{NC}$  is a common term to all flavor states, and does not affect the probability. It can be rotated away.
Thus
\begin{align}\label{eq:components}
    \bra{\nu_\b}H\ket{\nu_\a (x)} &= \sum_\eta\sum_k \left(U_{\b k} \frac{\dm_{k1}}{2E} U^*_{\eta k} + V_{CC}\delta_{\b\eta}\delta_{\eta e}\right) \psi_{\a\eta}(x) \nonumber \\
                                  &= i \dv{}{x}\psi_{\a\b}(x)\,.
\end{align}
If we form the vector 
\begin{align}
    \Psi_\a = \begin{pmatrix}
        \psi_{\a e} \\
        \psi_{\a \mu} \\
        \psi_{\a \tau}
    \end{pmatrix}\,,
\end{align}
we can write the Schrödinger equation on matrix form ($i \dv{}{x}\Psi_\a = H_F \Psi_\a$) and compare it with Eq.~\ref{eq:components} to see that the flavor Hamiltonian takes the form 
\begin{align}\label{eq:H_3gen}
    H_F &= \frac{1}{2E}(U M^2 U^\dagger + A) \nonumber \\
        &= \frac{1}{2E}\left[U \begin{pmatrix}
            0 & 0 & 0 \\
            0 & \dm_{21} & 0 \\
            0 & 0 & \dm_{31}
        \end{pmatrix} U^\dagger\right] + \sqrt{2}G_F N_e \begin{pmatrix}
            1 & 0 & 0 \\
            0 & 0 & 0 \\
            0 & 0 & 0
        \end{pmatrix}\,. 
\end{align}
This is the three-flavor neutrino oscillation Hamiltonian that we will solve numerically to obtain the evolution of $\Psi_\a$, whose squared components
are the probabilities
\begin{align}
    P_\alpha = \abs{\Psi_a}^2 &=\begin{pmatrix}
        \abs{\psi_{\a e}}^2 \\
        \abs{\psi_{\a \mu}}^{2} \\
        \abs{\psi_{\a \tau}}^{2}
    \end{pmatrix} \nonumber \\
    &=\begin{pmatrix}
        P_{\a e} \\
        P_{\a \mu} \\
        P_{\a \tau}
    \end{pmatrix} 
\end{align}

@@

\subsection{Two Neutrino Picture Solution}
By working in the limit where one of the neutrino masses can be neglected, 
we can reduce the complexity of the neutrino oscillations and obtain tractable analytical 
expressions. 
We now only consider the mixing between two neutrino flavors, say between $\ne, \nm$, and $\nu_1, \nu_2$. In other words, we let $\dm[31] = 0$. For simplicity, let $\delta_{CP} = 0$.
